Leaf traits and size drive differences in above-ground biomass distribution among woody plant functional types

### 2016-02-01 manuscript.Rnw
discussion work

Climate variables consistently added little explanatory power to the models when added to a model that included PFT and height (1 - 2.8 \%) (Table 1). Since this result is in contrast with recent finding of \citep{reich2014}, we also analyzed climate effects with methods similar to that study. When we tested for MAT effects on $M_F/M_S$ when only considering gymnosperms and angiosperms (thus grouping decidous and evergreen species), we found a significant and consistent positive relationship with MAT (Fig. S6) for angiosperms for most size classes (but never for gymnosperms). We found, however, that this was largely due to the fact that at higher MAT, an increasing proportion of angiosperms are evergreen (Fig. 1), which have higher $M_F/M_S$. When analyzed separately for deciduous and evergreen angiosperms, few and weaker significant relationships were found (Fig. S7).

@@

\begin{itemize}
\item MF/MS differed strongly between PFTs and was proportional to leaf mass per area, and steeply declined with plant height as expected.
\item We found that biomass distribution between leaves and stems was related to climate, but not when plant height and PFT - major confounding variables - were first accounted for. when PFT and height were first accounted for, we found weak and inconsistent effects of climate on all variables studied, with some exception for LMA. However, climate affects plant height, so we do expect global patterns in biomass distribution (Klein 2015, Simard 2011). Climate also affects, to some extent, distribution of major plant functional types (Fig. 1), explaining how our results are consistent with Reich et al. (2014), who reported shifts in biomass distribution with MAT. We found that for angiosperms, MF/MS increased with MAT, after splitting the dataset into five size classes This The significant relationship between MF/MS and MAT for angiosperms can be understood by the consistently higher proportion of evergreen angiosperms at higher MAT (Fig. 1), which, as a resutlt of higher LMA (Fig. 2) also have a higher MF/MS. The main conclusions from the climate analysis are certainly that PFT and plant height are by far the most important factors, even explaining most climate effects on biomass distribution.
\item If biomass distribution is related to leaf mass per area, an important trait, perhaps this can help to place plants on a continuous scale rather than rely on fixed categories. This is an active area of research, but is yet to be extended to patterns in biomass distribution.
\item Still lots and lots of variation to explain within PFTs; we do not argue for convergence, but rather that as a first step, we can understand differences between PFTs.
\item Leaf area was just about proportional to AS, except for the largest trees. This decreases because of the formation of heartwood (and AF/sapwood may decrease a little with height, but AF/AS more because of crown rise). Note that we have many small plants in the dataset as well, and within individual datasets it is more likely to pick up small-scale effects like these. 
\item Stem mass index: Chave biomass estimation. We find support for this, with MS/AS nearly proportional to height. Conifers diverged, for the subset of larger trees only. Not sure why these differences don't show up for small plants but variance was very high!
\item Some paragraph on future work, for example reconcile Poorter2015 with Reich2014 with our data.
\item Consequences for GVMs.
\end{itemize}

% We found that PFTs differed in LMF, and that LMF was proportional to LMA. This conclusion was particularly strong across the three PFTs studied, but it also held across species within PFT. The implication is that the amount of leaf area supported per unit biomass (leaf area ratio, LAR) does not differ between PFTs. This result seems robust, as our data include individual plants spanning the entire size range of woody plants in natural forests (0.1-100m) (Fig.~\ref{fig:figlmf}a). Previous studies have demonstrated a large difference in LMF between angiosperms and gymnosperms \citep{poorter2012,poorter2015}, but have not been able to explain these differences in terms of leaf traits. We show that LMA -- a central trait of the leaf economics spectrum (LES) \citep{wrightetal2004} -- explains differences in LMF in a consistent manner. 
% 
% We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012,poorter2015}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale, consistent with recent other results \citep{poorter2015}. 
% 
% The finding that LMF is proportional to LMA across PFTs implies that LAR does not differ systematically between PFTs. One explanation for this pattern lies in the strong positive correlation between LMA and leaf lifespan (LL) \citep{wrightetal2004}. Both \cite{poorter2012} and \citet{enquistniklas2002} explained higher LMF in gymnosperms by higher LL; plants simply maintain more cohorts of foliage, but this explanation demands that all else is equal. Yet it is not self-evident that everything else is indeed equal, that is, annual production of foliage could differ between PFTs. Nonetheless, further supporting this explanation, \citet{reich1992} showed a positive correlation between LL and total stand leaf biomass across diverse forest stands. Another, not mutually exclusive, possible explanation is that plants may allocate biomass in a manner that targets a more constant LAR rather than LMF. Canopy size scales with $M_T$ \citep{duursma2010}, and a larger canopy means higher light interception per unit total biomass \citep{duursma2007}. It can thus be argued that an optimal LAR exists that balances investment in supporting woody biomass (reflected by $M_T$) and the efficiency of foliage in terms of light interception. 
% 
% We found that patterns in $M_F/A_S$ across PFTs mirrored those of LMF (Fig.~\ref{fig:figlmf}). Because LMF may be decomposed into the terms $M_F/A_S$ and $M_T/A_S$ (eq. \ref{eq:M2}), this result suggests that $M_T/A_S$ does not vary between PFTs (see also Fig. S3). If this result holds, it means that aboveground biomass may be estimated across PFTs with a single equation that does not differ between PFTs. For tropical forests, \citet{chave2005} showed that aboveground biomass can be estimated as,
% 
% $$M_T = F \rho A_S H_T$$
% 
% where $F$ is a form factor, and $\rho$ wood density (kg m$^3$). After rearranging, this equation predicts that $M_T/A_S$ is proportional to plant height. We found some support for this prediction (Fig. S3), suggesting further improvements could be made by incorporating this dependence. It is timely to use database made public by \citet{falster2015} (BAAD) and \citet{poorter2015} to develop more general equations to estimate individual plant aboveground biomass across the globe.
% 
% We found that climate did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for this difference is that our measurements were taken on individual plants, whereas in  \citet{reich2014} they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig.~\ref{fig:fighist}), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Table \ref{table:statstable}).
% 
% Our results indicate that it is possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. Indeed, nearly as much variation was explained across species when using LMA instead of PFT in a statistical model. This is surprising because LMA only captures one aspect of functional differentiation among PFT -- leaf morphology (but including correlated effects on leaf lifespan and photosynthetic capacity via the leaf economics spectrum \citep{wrightetal2004}). We thus show that in a PFT-based classification, LMA is a good first estimate of biomass partitioning, however, more variation can be explained by another trait ($A_F/A_S$), which varies appreciably between species. These results suggest promising avenues for parameterising and simplifying biomass partitioning routines in GVMs.

@@

>>>>

Climate variables consistently added little explanatory power to the models when added to a model that included PFT and height (1 - 2.8 \%) (Table~\ref{table:varparttable}). Since this result is in contrast with recent finding of \citep{reich2014}, we also analyzed climate effects with methods similar to that study. When we tested for MAT effects on $M_F/M_S$ when only considering gymnosperms and angiosperms (thus grouping decidous and evergreen species), we found a significant and consistent positive relationship with MAT (Fig. S6) for angiosperms for most size classes (but never for gymnosperms). We found, however, that this was largely due to the fact that at higher MAT, an increasing proportion of angiosperms are evergreen (Fig. 1), which have higher $M_F/M_S$. When analyzed separately for deciduous and evergreen angiosperms, few and weaker significant relationships were found (Fig. S7).

@@

\begin{itemize}
\item MF/MS differed strongly between PFTs and was proportional to leaf mass per area, and steeply declined with plant height as expected.
\item We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012,poorter2015}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale, consistent with recent other results \citep{poorter2015}. 
\item We found that biomass distribution between leaves and stems was related to climate, but not when plant height and PFT - major confounding variables - were first accounted for. when PFT and height were first accounted for, we found weak and inconsistent effects of climate on all variables studied, with some exception for LMA. However, climate affects plant height, so we do expect global patterns in actual biomass distribution (Klein 2015, Simard 2011). Climate also affects, to some extent, distribution of major plant functional types (Fig. 1), explaining how our results are consistent with Reich et al. (2014), who reported shifts in biomass distribution with MAT. We found that for angiosperms, MF/MS increased with MAT, after splitting the dataset into five size classes This The significant relationship between MF/MS and MAT for angiosperms can be understood by the consistently higher proportion of evergreen angiosperms at higher MAT (Fig. 1), which, as a resutlt of higher LMA (Fig. 2) also have a higher MF/MS. However, based on a re-analysis of data in Reich et al. 2014 (not shown), this does not explain the pattern observed by them as it still held for gymnosperms and deciduous angiosperms. Whatever explains that different findings between us and Reich 2014, the main conclusions from the climate analysis are certainly that PFT and plant height are by far the most important factors, even explaining most climate effects on biomass distribution.
\item Still lots and lots of variation to explain within PFTs; we do not argue for convergence, but rather that as a first step, we can understand differences between PFTs. This helps parameterize PFTs, and also perhaps this can help to place plants on a continuous scale rather than rely on fixed categories. This is an active area of research, but is yet to be extended to patterns in biomass distribution. Our work contributes a bit to this.
\item Leaf area was just about proportional to AS, except for the largest trees. This decreases because of the formation of heartwood (and AF/sapwood may decrease a little with height, but AF/AS more because of crown rise). Note that we have many small plants in the dataset as well, and within individual datasets it is more likely to pick up small-scale effects like these. 
\item Stem mass index: Chave biomass estimation. We find support for this, with MS/AS nearly proportional to height. Conifers diverged, for the subset of larger trees only. Not sure why these differences don't show up for small plants but variance was very high! For tropical forests, \citet{chave2005} showed that aboveground biomass can be estimated as,

$$M_T = F \rho A_S H_T$$

where $F$ is a form factor, and $\rho$ wood density (kg m$^3$). This explains why (for larger trees) at a given AS, conifers had lower stem mass (Fig. 5).
\item Consequences for GVMs.

\end{itemize}

@@

\begin{itemize}
\item Some paragraph on future work, for example reconcile Poorter2015 with Reich2014 with our data.
\end{itemize}
